---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Loading required packages into R
```{r}
library(knitr)
library(ggplot2)
library(dplyr)
library(plyr)
```

# Reading the dataset and processing the data
activity.csv and checking first part of file
```{r}
activity <- read.csv("activity.csv")

totalstepsperday <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
head(totalstepsperday)
```

converting dates to Y-M-D format
```{r}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

# Creating a histogram with 'total number of steps'
calculate steps as it relates to date using SUM (per day)
```{r}
hist(totalstepsperday$steps, 
     main="Total Steps per Day", 
     xlab="Number of Steps per Day", 
     ylab = "Interval",
     col="orange",
     breaks=50)
```

# Calculate Mean and Median number of steps taken each day
mean of total steps per day
```{r}
msteps <- mean(totalstepsperday$steps)
msteps
```

median of total steps per day
```{r}
medsteps <- median(totalstepsperday$steps)
medsteps
```

check work using summary
```{r}
summary(totalstepsperday)
```

# Create time-series plot for average number of steps taken
five minute average using steps to interval - FUN = mean instead of sum
```{r}
fivemin <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
```

line chart
```{r}
plot(x = fivemin$interval, 
     y = fivemin$steps, 
     type = "l", 
     col = "orange",
     xlab = "5-minute Intervals",
     ylab = "Average Steps Taken ~ Days",
     main = "Average Daily Activity Pattern")
```
Find 5-minute interval with most steps taken
```{r}
maxsteps <- fivemin$interval[which.max(fivemin$steps)]
maxsteps
```
# Strategy for imputing missing data
Using mean value to replace NA
```{r}
activity2 <- activity
nas <- is.na(activity2$steps)
avg_interval <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify = TRUE)
activity2$steps[nas] <- avg_interval[as.character(activity2$interval[nas])]
names(activity2)
```

Check for no-NA
```{r}
sum(is.na(activity2))
```


# Create histogram for no-NA data
Setting up the pannel for one row and two columns
```{r}
par(mfrow=c(1,2))
```

Similar analysis without NAs now
```{r}
totalstepsperday2 <- aggregate(steps ~ date, data = activity2, FUN = sum, na.rm = TRUE)
head(totalstepsperday2)
```

Histogram without the NA values
```{r}
hist(totalstepsperday2$steps, 
     main = "Total Steps per Day (no-NA)", 
     xlab = "Number of Steps per Day", 
     ylab = "Interval",
     col="green",
     breaks=50)
```

Histogram with the orginal dataset
```{r}
hist(totalstepsperday$steps, 
     main="Total Steps per Day (Original)", 
     xlab="Number of Steps per Day", 
     ylab = "Interval",
     col="orange",
     breaks=50)
```

Resetting the panel
```{r}
par(mfrow=c(1,1))
```

What is the impact of imputing data?
```{r}
summary(totalstepsperday)

summary(totalstepsperday2)
```

# Create panel plot to compare steps during the week or weekend
Add the new weekend/weekday field
```{r}
activity2<- activity2%>%
        mutate(typeofday= ifelse(weekdays(activity2$date)=="Saturday" | weekdays(activity2$date)=="Sunday", "Weekend", "Weekday"))
head(activity2)
```
Plot - Line chart
```{r}
fivemin2<- aggregate(steps ~ interval + typeofday, data = activity2, FUN = mean)
head(fivemin2)

g <- qplot(interval, steps, data = fivemin2, facets = typeofday~.)
g + geom_line(size = 1) + ylab("Mean steps") + ggtitle("Average number of steps taken per daytype")
```
